A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-05-10 paper/paper-jsdse/paper-jsdse.Rmd
update of intro according to AE comment

---
authors:
  - name: Dewi Amaliah
    thanks: "Corresponding author, dlamaleeah@gmail.com"
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dlamaleeah@gmail.com
  - name: Dianne Cook
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dicook@monash.edu
  - name: Emi Tanaka
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: emi.tanaka@monash.edu
  - name: Kate Hyde
    thanks: ""
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: "hyde.kate.a@gmail.com"
  - name: Nicholas Tierney
    thanks: ""
    department: 
    affiliation: Telethon Kids Institute
    location: Nedlands, WA 6009
    email: "nicholas.tierney@gmail.com"
blind: 0 # 0 for no
bibliography: ../references.bib
#preamble: >
output: 
  bookdown::pdf_document2:
    template: TAS-template-for-rmd.tex
    keep_tex: true
date: "`r Sys.Date()`"
always_allow_html: true
---

>>>>

---
authors:
  - name: Dewi Amaliah
    thanks: "Corresponding author, dlamaleeah@gmail.com"
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dlamaleeah@gmail.com
  - name: Dianne Cook
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: dicook@monash.edu
  - name: Emi Tanaka
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: emi.tanaka@monash.edu
  - name: Kate Hyde
    thanks: ""
    department: Dept of Econometrics and Business Statistics
    affiliation: Monash University
    location: Clayton, VIC 3800
    email: "hyde.kate.a@gmail.com"
  - name: Nicholas Tierney
    thanks: ""
    department: "Geospatial Health and Development"
    affiliation: Telethon Kids Institute
    location: Nedlands, WA 6009
    email: "nicholas.tierney@gmail.com"
blind: 0 # 1 for yes
bibliography: ../references.bib
#preamble: >
output: 
  bookdown::pdf_document2:
    template: TAS-template-for-rmd.tex
    keep_tex: true
date: "`r Sys.Date()`"
always_allow_html: true
---

### 2022-05-10 paper/paper-jsdse/paper-jsdse.tex
update of intro according to AE comment

\documentclass[12pt]{article}
\usepackage{amsmath}
%\usepackage{longtable}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{tcolorbox}
\usepackage{fontawesome}
\usepackage{booktabs}
\usepackage{float}
\usepackage{url} % not crucial - just used below for the URL
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{1}

@@

\if0\blind
{
  \title{\bf A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database}
  \author{
            Dewi Amaliah\thanks{Corresponding author, \href{mailto:dlamaleeah@gmail.com}{\nolinkurl{dlamaleeah@gmail.com}}}\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Dianne Cook\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Emi Tanaka\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Kate Hyde\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Nicholas Tierney\\
      , Telethon Kids Institute, }
        \maketitle
} \fi

@@

The story from modeling the data (and as reported by Singer and Willett) is that wages increase with the length of time in the workforce, a higher level of education leads to higher wages, and that race makes a difference, on average. However, an exploratory analysis reveals that the individual experience varies a great deal from the overall average. Some individuals experience a decline in wages the longer they are in the workforce, and many experience volatility in their wages. The wages data was used to illustrate exploratory longitudinal data analysis in Ilk (2004), and was further developed into a case study for use in the teaching of exploratory data analysis at Iowa State University.

This disparity between the average and the individual is a part of statistics as a discipline that requires more attention. This particular data set is a prime example of discussing this disparity. Textbook data sets have longevity if there is an unresolved mystery. The iris data (Andereson 1935) is a prime example. It has withstood the test of time because the three species cannot be perfectly classified, and so it continues to challenge researchers and instructors to do better in the analysis (a side note: the iris data is best replaced today with the penguins data (Horst, Hill, and Gorman 2020), which has similar qualities, is new, and does not suffer from a connection with eugenics (Stodel 2020)). We argue that the wages data is in this class of textbook data, too, because it presents a challenge for longitudinal data analysis: how can we better summarize and explain the individual experience?

For statistics and data science by association, it is increasingly important to reach the individual. One might describe this as a divergence of purpose -- statistics for public policy -- or statistics for the public. The two are not the same. As the world becomes more electronically connected, combating misinformation and mitigating conspiracy theories require statistics that address the individual. For example, with the wages data, even though the message for public policy is that demographic profile is related to different wage patterns on average, the message for the individual is you are more than your demographic. The majority of people in the study do not have a pattern that is similar to the average. If you have a bad experience, that your wages have declined over time, you are not alone; there are others like you, and more than you think. A similar tone is echoed occasionally in the public media. For example, an article published in the Sydney Morning Herald argues there is no average Australian (Moncrief 2015).

As a textbook data set, though, the wages data is outdated. The most recent year in the data is 1994, 10 years prior to when Singer and Willett (2003) was published. Teachers of statistics need contemporary data sets to show how techniques are relevant for today's students. Using tired old textbook data sets can imbue a misconception that the field is not current. The wages data is extracted from NSLY79, one of the best examples of open data, which is constantly being updated. It should be possible to continuously refresh the textbook data from the data repository. This paper describes our (non-glamorous) journey from open (Open Knowledge Foundation 2021) wild data to textbook data.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-iris-data}{}}%
Andereson, Edgar. 1935. {``The Irises of the Gaspe Peninsula.''} \emph{Bulletin of the American Iris Society} 59: 2--5.

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
%\usepackage{longtable}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{tcolorbox}
\usepackage{fontawesome}
\usepackage{booktabs}
%\usepackage{float}
\usepackage{url} % not crucial - just used below for the URL
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

@@

\if0\blind
{
  \title{\bf A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database}
  \author{
            Dewi Amaliah\thanks{Corresponding author, \href{mailto:dlamaleeah@gmail.com}{\nolinkurl{dlamaleeah@gmail.com}}}\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Dianne Cook\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Emi Tanaka\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Kate Hyde\\
      Dept of Econometrics and Business Statistics, Monash University, \\
            Nicholas Tierney\\
      Geospatial Health and Development, Telethon Kids Institute, }
        \maketitle
} \fi

@@

The story from modeling the data (and as reported by Singer and Willett) is that wages increase with the length of time in the workforce, a higher level of education leads to higher wages, and that race makes a difference, on average. However, an exploratory analysis reveals that the individual experience varies a great deal from the overall average. Some individuals experience a decline in wages the longer they are in the workforce, and many experience volatility in their wages. It is for these reasons that the wages data was used to illustrate exploratory longitudinal data analysis in Ilk (2004), and was further developed into a case study for use in the teaching of exploratory data analysis at Iowa State University.

This disparity between the overall trend and the individual heterogeneity is what makes this data interesting. Textbook data sets have longevity if they have unresolved elements. The iris data (Anderson 1935) is a prime example. It has withstood the test of time because the three species cannot be perfectly classified, and so it continues to challenge researchers and instructors to do better in the analysis. (It should be noted that the iris data is best replaced today with the penguins data (Horst, Hill, and Gorman 2020), which has similar qualities, is new, and does not suffer from a connection with eugenics (Stodel 2020)). We argue that the wages data is in this class of textbook data, too, because it presents a challenge for longitudinal data analysis: how can we better summarize and explain the individual experience?

For statistics, and data science by association, it is increasingly important to reach the individual. One might think of this as a divergence of purpose -- statistics for public policy -- or statistics for the public. The two are not the same. As the world becomes more electronically connected, combating misinformation and mitigating conspiracy theories require statistics that address the individual. For example, with the wages data, the overall trend, across demographics, is a steady increase in wages, but the patterns among individuals is heterogeneous. Some individuals see a decline in their wages, some an increase, some have dramatic ups and downs from year to year, and the heterogeneity of the patterns are evident across demographics. Subject-specific variation explains most of the variation in wages, and even though the overall trend is statistically significant, it is weak. The message for public policy is that demographic profile is related to different wage patterns on average and some structural social change is desirable. However, discussing the overall trend with an individual is misleading -- more than likely their pattern is completely different. It could even be depressing for them to compare themselves to the overall average, especially if the individual's wages have a declining trend or high volatility. It would be more useful for an individual to know what percentage of the population have profiles like theirs. We argue that different summaries are needed to communicate with the public, and ones where they might be able to individually act on to change their own situation. We occasionally hear this voiced in the public media media, also, for example, an article published in the Sydney Morning Herald argues there is no average Australian (Moncrief 2015). Thus the wages data provides educators with a challenge, from a methodological and a philosophical perspective, and wage experience is a topic of interest to many.

As a textbook data set, though, the wages data is outdated. The most recent year in the data is 1994, 10 years prior to when Singer and Willett (2003) was published. Teachers of statistics need contemporary data sets to show how techniques are relevant for today's students. Using tired old textbook data sets can imbue a misconception that the field is not current. The wages data is extracted from NSLY79, one of the best examples of open data (see details at Open Knowledge Foundation (2021)), which is constantly being updated. It should be possible to continuously refresh the textbook data from the data repository. This paper describes our (non-glamorous) journey from open wild data to textbook data.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-iris-data}{}}%
Anderson, Edgar. 1935. {``The Irises of the Gaspe Peninsula.''} \emph{Bulletin of the American Iris Society} 59: 2--5.

### 2022-05-10 paper/paper.Rmd
update of intro according to AE comment

The story from modeling the data (and as reported by Singer and Willett) is that wages increase with the length of time in the workforce, a higher level of education leads to higher wages, and that race makes a difference, on average. However, an exploratory analysis reveals that the individual experience varies a great deal from the overall average. Some individuals experience a decline in wages the longer they are in the workforce, and many experience volatility in their wages. The wages data was used to illustrate exploratory longitudinal data analysis in @ilk2004, and was further developed into a case study for use in the teaching of exploratory data analysis at `r censor_word("Iowa State University")`.

@@

This disparity between the average and the individual is a part of statistics as a discipline that requires more attention. This particular data set is a prime example of discussing this disparity. Textbook data sets have longevity if there is an unresolved mystery. The iris data [@iris-data] is a prime example. It has withstood the test of time because the three species cannot be perfectly classified, and so it continues to challenge researchers and instructors to do better in the analysis (a side note: the iris data is best replaced today with the penguins data [@penguins-data], which has similar qualities, is new, and does not suffer from a connection with eugenics [@notiris]). We argue that the wages data is in this class of textbook data, too, because it presents a challenge for longitudinal data analysis: how can we better summarize and explain the individual experience?

For statistics and data science by association, it is increasingly important to reach the individual. One might describe this as a divergence of purpose -- statistics for public policy -- or statistics for the public. The two are not the same. As the world becomes more electronically connected, combating misinformation and mitigating conspiracy theories require statistics that address the individual. For example, with the wages data, even though the message for public policy is that demographic profile is related to different wage patterns on average, the message for the individual is you are more than your demographic. The majority of people in the study do not have a pattern that is similar to the average. If you have a bad experience, that your wages have declined over time, you are not alone; there are others like you, and more than you think. A similar tone is echoed occasionally in the public media. For example, an article published in the Sydney Morning Herald argues there is no average Australian [@notaverage].

As a textbook data set, though, the wages data is outdated. The most recent year in the data is 1994, 10 years prior to when @SingerJudithD2003Alda was published. Teachers of statistics need contemporary data sets to show how techniques are relevant for today's students. Using tired old textbook data sets can imbue a misconception that the field is not current. The wages data is extracted from NSLY79, one of the best examples of open data, which is constantly being updated. It should be possible to continuously refresh the textbook data from the data repository. This paper describes our (non-glamorous) journey from open [@opendata] wild data to textbook data.

>>>>

The story from modeling the data (and as reported by Singer and Willett) is that wages increase with the length of time in the workforce, a higher level of education leads to higher wages, and that race makes a difference, on average. However, an exploratory analysis reveals that the individual experience varies a great deal from the overall average. Some individuals experience a decline in wages the longer they are in the workforce, and many experience volatility in their wages. It is for these reasons that the wages data was used to illustrate exploratory longitudinal data analysis in @ilk2004, and was further developed into a case study for use in the teaching of exploratory data analysis at `r censor_word("Iowa State University")`.

@@

This disparity between the overall trend and the individual heterogeneity is what makes this data interesting. Textbook data sets have longevity if they have unresolved elements. The iris data [@iris-data] is a prime example. It has withstood the test of time because the three species cannot be perfectly classified, and so it continues to challenge researchers and instructors to do better in the analysis. (It should be noted that the iris data is best replaced today with the penguins data [@penguins-data], which has similar qualities, is new, and does not suffer from a connection with eugenics [@notiris]). We argue that the wages data is in this class of textbook data, too, because it presents a challenge for longitudinal data analysis: how can we better summarize and explain the individual experience?

For statistics, and data science by association, it is increasingly important to reach the individual. One might think of this as a divergence of purpose -- statistics for public policy -- or statistics for the public. The two are not the same. As the world becomes more electronically connected, combating misinformation and mitigating conspiracy theories require statistics that address the individual. For example, with the wages data, the overall trend, across demographics, is a steady increase in wages, but the patterns among individuals is heterogeneous. Some individuals see a decline in their wages, some an increase, some have dramatic ups and downs from year to year, and the heterogeneity of the patterns are evident across demographics. Subject-specific variation explains most of the variation in wages, and even though the overall trend is statistically significant, it is weak. The message for public policy is that demographic profile is related to different wage patterns on average and some structural social change is desirable. However, discussing the overall trend with an individual is misleading -- more than likely their pattern is completely different. It could even be depressing for them to compare themselves to the overall average, especially if the individual's wages have a declining trend or high volatility. It would be more useful for an individual to know what percentage of the population have profiles like theirs. We argue that different summaries are needed to communicate with the public, and ones where they might be able to individually act on to change their own situation. <!--this information is the appropriate message for an individual is that you as an individual are far more than your demographic. The majority of people in the study do not have a pattern that is similar to the overall trend. If you have experienced declining wages, you are not alone; there are others like you, and more than you think.--> We occasionally hear this voiced in the public media media, also, for example, an article published in the Sydney Morning Herald argues there is no average Australian [@notaverage]. Thus the wages data provides educators with a challenge, from a methodological and a philosophical perspective, and wage experience is a topic of interest to many.

As a textbook data set, though, the wages data is outdated. The most recent year in the data is 1994, 10 years prior to when @SingerJudithD2003Alda was published. Teachers of statistics need contemporary data sets to show how techniques are relevant for today's students. Using tired old textbook data sets can imbue a misconception that the field is not current. The wages data is extracted from NSLY79, one of the best examples of open data (see details at @opendata), which is constantly being updated. It should be possible to continuously refresh the textbook data from the data repository. This paper describes our (non-glamorous) journey from open wild data to textbook data.

### 2022-05-10 paper/paper.tex
update of intro according to AE comment

\begin{abstract}
Textbook data is essential for teaching statistics and data science methods because they are clean, allowing the instructor to focus on methodology. Ideally textbook data sets are refreshed regularly, especially when they are subsets taken from an on-going data collection. It is also important to use contemporary data for teaching, to imbue the sense that the methodology is relevant today. This paper describes the trials and tribulations of refreshing a textbook data set on wages, extracted from the National Longitudinal Survey of Youth (NLSY79) in the early 1990s. The data is useful for teaching modeling and exploratory analysis of longitudinal data. Subsets of NLSY79, including the wages data, can be found in supplementary files from numerous textbooks and research articles. The NLSY79 database has been continuously updated through to 2018, so new records are available. Here we describe our journey to re-create the wages data, and document the process so that the data can be regularly updated into the future. Our journey was difficult because the steps and decisions taken to get from the raw data to the wages textbook subset have not been clearly articulated. We have been diligent to provide a reproducible workflow for others to follow, which also hopefully inspires more attempts at refreshing data for teaching. Three new data sets and the code to produce them are provided in the open source R package called {[}CENSORED{]}.
\end{abstract}

@@

For statistics and data science by association, it is increasingly important to reach the individual. One might describe this as a divergence of purpose -- statistics for public policy -- or statistics for the public. The two are not the same. As the world becomes more electronically connected, combating misinformation and mitigating conspiracy theories require statistics that address the individual. For example, with the wages data, even though the message for public policy is that demographic profile is related to different wage patterns on average, the message for the individual is you are more than your demographic. The majority of people in the study do not have a pattern that is similar to the average. If you have a bad experience, that your wages have declined over time, you are not alone; there are others like you, and more than you think. A similar tone is echoed occasionally in the public media. For example, an article published in the Sydney Morning Herald argues there is no average Australian (Moncrief 2015).

@@

Figure \ref{fig:compare-plot} shows the mean hourly wage before and after the extreme values are replaced. The plot shows fluctuations in wages remain, but the large spikes (in this sample, individuals 8296, 9962), which are considered implausible, are replaced. The \texttt{mean\_hourly\_wage} and \texttt{year} are set as the dependent and predictor, respectively. Furthermore, we use iteratively reweighted least squares with Huber weighting, where the observation with a slight residual gets a weight of 1, while the larger the residual, the smaller the weight (less than 1) (UCLA: Statistical Consulting Group 2021).

The challenging part of detecting an anomaly using the robustness weight is determining the weight threshold beyond which the observations are considered outliers. It is also important not to be too aggressive in outlier removal, because this would overly smooth the data, and ignore actual workforce experiences important wage volatility such as, abnormally high wages for a fleeting amount of time. To explore the risk of being overly vigorous in labeling observations as outliers, some testing of threshold value was done. Samples of individuals were examined under the different thresholds with an eye to the smoothness of the profiles. Overly smooth profiles would indicate the replacement of values was too severe, resulting in the removing the very interesting volatility of wages seen in many individuals. A threshold of 0.12 was chosen. That struck a balance between maintaining the natural variability of the wages with minimizing implausible values.
Using this threshold, we impute the observations whose weights are less than 0.12 with the models' predicted value. We then flag those observations in a new variable called \texttt{is\_pred}, so that this change can be monitored in the downstream analyses.

@@

{\centering \includegraphics[width=0.85\linewidth]{/Users/cookd/numbats/yowie-paper/paper/figures/flowchart_blind}

@@

This paper has illustrated the steps and decisions made to take a particular open data set and make it a textbook data set, ready for the classroom or research. In the first stage, we showed the steps performed to get the data from the NLSY79 database. The data format was converted to a tidy format for more flexibility in cleaning and exploring. Initial data analysis was conducted to investigate and screen the quality of the data. We found and provided a fix for many anomalous observations in wages using a robust linear regression model. The refreshed data is compared with the original set using a variety of numerical summaries and graphics. The current subset is made available in a new R package, called \texttt{[CENSORED]}.

The data cleaning process is documented, and the code has been made available. These provide the opportunity to again refresh the textbook data as new data is published into the NLSY79 database.
Determining an appropriate robustness weight from which to threshold unusual observations was conducted using a \texttt{shiny} (Chang et al. 2020) app, and the choices used in the refreshed data are documented.

@@

\leavevmode\vadjust pre{\hypertarget{ref-rlm}{}}%
UCLA: Statistical Consulting Group. 2021. {``{Robust Regression \textbar{} R Data Analysis Examples}.''} February 2021. \url{https://stats.idre.ucla.edu/r/dae/robust-regression/}.

>>>>

\begin{abstract}
Textbook data is essential for teaching statistics and data science methods because they are clean, allowing the instructor to focus on methodology. Ideally textbook data sets are refreshed regularly, especially when they are subsets taken from an on-going data collection. It is also important to use contemporary data for teaching, to imbue the sense that the methodology is relevant today. This paper describes the trials and tribulations of refreshing a textbook data set on wages, extracted from the National Longitudinal Survey of Youth (NLSY79) in the early 1990s. The data is useful for teaching modeling and exploratory analysis of longitudinal data. Subsets of NLSY79, including the wages data, can be found in supplementary files from numerous textbooks and research articles. The NLSY79 database has been continuously updated through to 2018, so new records are available. Here we describe our journey to re-create the wages data, and document the process so that the data can be regularly updated into the future. Our journey was difficult because the steps and decisions taken to get from the raw data to the wages textbook subset have not been clearly articulated. We have been diligent to provide a reproducible workflow for others to follow, which also hopefully inspires more attempts at refreshing data for teaching. Three new data sets and the code to produce them are provided in the open source R package called \texttt{yowie}.
\end{abstract}

@@

For statistics, and data science by association, it is increasingly important to reach the individual. One might describe this as a divergence of purpose -- statistics for public policy -- or statistics for the public. The two are not the same. As the world becomes more electronically connected, combating misinformation and mitigating conspiracy theories require statistics that address the individual. For example, with the wages data, the overall trend, across demographics, is a steady increase in wages, but the patterns among individuals is heterogeneous. Some individuals see a decline in their wages, some an increase, some have dramatic ups and downs from year to year. The heterogeneous patterns occur across demographics. The subject-specific variation explains most of the variation in wages, even though the overall trend is statistically significant, it is weak. The the message for public policy is that demographic profile is related to different wage patterns on average and some structural social change is desirable. However, this information is the appropriate message for an individual is that you as an individual are far more than your demographic. The majority of people in the study do not have a pattern that is similar to the overall trend. If you have experienced declining wages, you are not alone; there are others like you, and more than you think. A similar tone is echoed occasionally in the public media. For example, an article published in the Sydney Morning Herald argues there is no average Australian (Moncrief 2015).

@@

Figure \ref{fig:compare-plot} shows the profiles for a sample of 30 individuals the mean hourly wage before and after extreme values are replaced. The plot shows fluctuations in wages remain, but the large spikes (in this sample, individuals 8296, 9962), which are considered implausible, are replaced.

The challenging part of detecting an anomaly using the robustness weight is determining the weight threshold beyond which the observations are considered outliers. It is also important not to be too aggressive in outlier removal, because this would overly smooth the data, and ignore actual workforce experiences important wage volatility such as, abnormally high wages for a fleeting amount of time. To explore the risk of being overly vigorous in labeling observations as outliers, a range of threshold value were assessed. Numerous samples of individuals were drawn for each threshold, and the smoothness of the profiles was examined. Overly smooth profiles would indicate the replacement of values was too severe, resulting in the removing the very interesting volatility of wages seen in many individuals. This was achieved using a purpose written \texttt{shiny} (Chang et al. 2020) app (provided with the code of this paper). Observations where a change was made are tracked with a new variable called \texttt{is\_pred}, so that this effect in the downstream analyses can be monitored.

@@

{\centering \includegraphics[width=0.85\linewidth]{figures/flowchart_blind}

@@

This paper has illustrated the steps and decisions made to take a particular open data set and make it a textbook data set, ready for the classroom or research. In the first stage, we showed the steps performed to get the data from the NLSY79 database. The data format was converted to a tidy format for more flexibility in cleaning and exploring. Initial data analysis was conducted to investigate and screen the quality of the data. We found and provided a fix for many anomalous observations in wages using a robust linear regression model. The refreshed data is compared with the original set using a variety of numerical summaries and graphics. The current subset is made available in a new R package, called \texttt{yowie}.

The data cleaning process is documented, and the code has been made available. These provide the opportunity to again refresh the textbook data as new data is published into the NLSY79 database. Determining an appropriate robustness weight from which to threshold unusual observations was conducted using a \texttt{shiny} (Chang et al. 2020) app, which is provided with the code, and the choices used in the refreshed data are documented.

@@
